\texttt{Davos}: a Python package ``smuggler'' for constructing lightweight reproducible notebooks

### 2023-10-26 paper/main.tex
additional edits made with PCF based on our in-person discussion.  also reverted fig 1 + last sentence of fig 1 caption per our conversation.

\begin{figure}[tp] \centering
\includegraphics[width=0.6\textwidth]{figs/shareable_code_2d} \caption{\small
\textbf{Systems for ensuring code reproducibility within the Python ecosystem.}
The $x$-axis denotes the ``burden'' placed on users to install and configure
the given system (systems placed further to the right, that fall within the
redder shading, impose a higher setup cost on the user). The $y$-axis denotes
the degree to which the system guarantees that the code will run similarly for
different users (systems placed higher up, that fall within the bluer shading,
offer stronger guarantees). From left to right, bottom to top: plain-text
\textbf{Python scripts} (\texttt{.py} files) provide the most basic ``system''
for sharing raw code. Scripts may reference external packages, but those
packages must be manually installed on other users' systems. Further, any
checking needed to verify that the correct versions of those packages were
installed must also be performed manually. \textbf{Jupyter notebooks}
(\texttt{.ipynb} files) comprise embedded text, executable code, and media
(including rendered figures, code output, etc.). When the \textbf{Davos
package} is imported into a Jupyter notebook, the notebook's functionality is
extended to automatically install any required external packages (at their
correct versions, when specified) into an isolated directory kept separate from
the user's existing set of packages. \textbf{Virtual environments} (along with
dependency specification files) similarly allow users to install required
packages into an isolated directory, and \textbf{Python environments} (again,
along with dependency specification files) extend this to include a specific
version of Python itself. However, both of these systems require users to
create, populate, and manage environments manually. This typically entails
distributing a configuration file (e.g., a \texttt{requirements.txt},
\texttt{pyproject.toml}~\cite{CannEtal16}, or \texttt{environment.yml} file)
that specifies all project dependencies (including version numbers) alongside
the primary code base. Users can then install a third-party
tool~\cite[e.g.,][]{BickEtal07, Eust19, Anac12} to read the file and build the
environment. \textbf{Containers} provide a means of defining an isolated
environment that includes a complete operating system (independent of the
user's operating system), in addition to (optionally) specifying a virtual
environment or other configurations needed to provide the necessary computing
environment. Containers are typically defined using specification files (e.g.,
a plain-text \texttt{Dockerfile}~\cite{Merk14}) that instruct the
virtualization engine regarding how to build the containerized environment.
\textbf{Virtual machines} provide a complete hardware-level simulation of the
computing environment. In addition to simulating specific hardware, virtual
machines (typically specified using binary image files) must also define
operating system-level properties of the computing environment. $\Delta_1$
represents the increased reproducibility \texttt{Davos} provides over standard
Jupyter notebooks for no greater setup cost. $\Delta_2$ represents
\texttt{Davos}'s lower setup cost compared to standard virtual environments.}
\label{fig:code-sharing} \end{figure}

@@

We designed \texttt{Davos} to occupy a ``sweet spot'' within this space.
\texttt{Davos} is a notebook-installable package that adds functionality to the
default notebook experience. Like standard Jupyter notebooks,
\texttt{Davos}-enhanced notebooks allow researchers to include text, executable
code, and media within a single file. No further setup or installation is
required from the user, beyond what is needed to run a standard Jupyter
notebook. And like virtual environments or Python environments (when these
tools are bundled with dependency specification files), \texttt{Davos} provides
a convenient mechanism for fully specifying (and installing, as needed) a
complete set of Python dependencies, including specific package versions, which
are contained and isolated from the rest of the user's system.

>>>>

\begin{figure}[tp] \centering
\includegraphics[width=0.6\textwidth]{figs/shareable_code_2d} \caption{\small
\textbf{Systems for ensuring code reproducibility within the Python ecosystem.}
The $x$-axis denotes the ``burden'' placed on users to install and configure
the given system (systems placed further to the right, that fall within the
redder shading, impose a higher setup cost on the user). The $y$-axis denotes
the degree to which the system guarantees that the code will run similarly for
different users (systems placed higher up, that fall within the bluer shading,
offer stronger guarantees). From left to right, bottom to top: plain-text
\textbf{Python scripts} (\texttt{.py} files) provide the most basic ``system''
for sharing raw code. Scripts may reference external packages, but those
packages must be manually installed on other users' systems. Further, any
checking needed to verify that the correct versions of those packages were
installed must also be performed manually. \textbf{Jupyter notebooks}
(\texttt{.ipynb} files) comprise embedded text, executable code, and media
(including rendered figures, code output, etc.). When the \textbf{Davos
package} is imported into a Jupyter notebook, the notebook's functionality is
extended to automatically install any required external packages (at their
correct versions, when specified) into an isolated directory kept separate from
the user's existing set of packages. \textbf{Virtual environments} (along with
dependency specification files) similarly allow users to install required
packages into an isolated directory, and \textbf{Python environments} (again,
along with dependency specification files) extend this to include a specific
version of Python itself. However, both of these systems require users to
create, populate, and manage environments manually. This typically entails
distributing a configuration file (e.g., a \texttt{requirements.txt},
\texttt{pyproject.toml}~\cite{CannEtal16}, or \texttt{environment.yml} file)
that specifies all project dependencies (including version numbers) alongside
the primary code base. Users can then install a third-party
tool~\cite[e.g.,][]{BickEtal07, Eust19, Anac12} to read the file and build the
environment. \textbf{Containers} provide a means of defining an isolated
environment that includes a complete operating system (independent of the
user's operating system), in addition to (optionally) specifying a virtual
environment or other configurations needed to provide the necessary computing
environment. Containers are typically defined using specification files (e.g.,
a plain-text \texttt{Dockerfile}~\cite{Merk14}) that instruct the
virtualization engine regarding how to build the containerized environment.
\textbf{Virtual machines} provide a complete hardware-level simulation of the
computing environment. In addition to simulating specific hardware, virtual
machines (typically specified using binary image files) must also define
operating system-level properties of the computing environment. $\Delta_1$
represents the increased reproducibility \texttt{Davos} provides over standard
Jupyter notebooks for no greater setup cost. $\Delta_2$ represents
\texttt{Davos}'s lower setup cost compared to standard virtual environments
despite more stringently ensuring reproducibility.} \label{fig:code-sharing}
\end{figure}

@@

We designed \texttt{Davos} to occupy a ``sweet spot'' within this space.
\texttt{Davos} is a notebook-installable package that adds functionality to the
default notebook experience. Like standard Jupyter notebooks,
\texttt{Davos}-enhanced notebooks allow researchers to include text, executable
code, and media within a single file. No further setup or installation is
required from the user, beyond what is needed to run a standard Jupyter
notebook. And like separate dependency specification files used to build
virtual environments or Python environments, \texttt{Davos} provides a
convenient mechanism for fully specifying (and installing, as needed) a
complete set of Python dependencies, including specific package versions, which
are contained and isolated from the rest of the user's system.
